\paragraph{Rui Pinto}
was born in October of 1988 and raised in \textbf{"Mafamude"}, a small town in the North of Portugal.
Since very early in his life, Pinto was very invested in football, he was especially fanatic for FC Porto, a football club established very close to his home town.

\textit{"Faculdade de Letras da Universidade do Porto"} is where he cultivated his education, where he studied History, and where he enrolled in the Erasmus program to study abroad in Budapest, Hungary.

While being abroad, in 2013, Pinto was involved in several cyber attacks to the \textit{Caledonian Bank} which had the objective of transferring \textbf{264} thousand dollars to the \textit{Deutsche Bank} but was ultimately discovered by the authorities and detained before the transaction was actually completed.

With help from his lawyer, Aníbal Pinto, he was saved from a bigger sentence that lead in 2014 to the decision from the \textit{"Departamento de Investigação e Ação Penal do Porto"} to archive the case and settle for an extralegal accord between Rui Pinto and the Bank, out of which resulted a \textbf{100} thousand dollar fine.
\footnote{[14-05-2022 14:25] \url{https://desporto.sapo.pt/geral/artigos/football-leaks-rui-pinto-apresenta-queixa-contra-portugal-na-uniao-europeia}}


\paragraph{Football Leaks} self declares itself as an attempt to expose the real nature of football and its organizations.
With these ambitions in mind, \url{http://football-leaks.livejournal.com} was created on the 29th of September 2015.
John was the name behind the publications on the blog, posting thousands of confidential documents, making public several contracts and transactions, which expose multiple tax evasion schemes committed around the world.

Until today, Rui Pinto is the only one that has publicly claimed ownership of the famous website.

\paragraph{The crimes} that Rui Pinto is accused of by the Public Ministry of Portugal on September of 2019, after being arrested by the Hungarian authorities (16th of January 2019) and brought back to Portugal tp be kept in preventive arrest, are 147 in total, 70 violations of correspondence, 75 of which involving illegitimate access, one of attempted extortion and one of computer sabotage.

Doyen Sports is the football investment fund that accused Football Leaks of persuading several big Clubs like Sporting, which in particular had the contract of his current coach, "Jorge Jesus", and documents related to the renovation and transaction of the peruvian player "André Carrilho" publicly posted on the alleged website.
\footnote{[15-05-2022 13:16] \url{https://observador.pt/2015/12/17/football-leaks-acusada-tentar-extorquir-dinheiro/}}

\paragraph{The european wistleblowers prize} was awarded to Rui Pinto while he was in prison, alongside Julian Assange and Yasmine Motarjemi on the 16th of April 2019.

\paragraph{The Cooperation} between Rui Pinto and the authorities from France, Belgium, Netherlands and even with "Polícia Judiciária" from Portugal has been very productive as confirmed by Luís Neves (Director at "Polícia Judiciária") in court, he stated: \textit{There is that effective collaboration...}. and \textit{...many disks had not been opened}.
\footnote{ [15-05-2022 16:19] translated from \textit{"Há essa colaboração efetiva"} and \textit{"muitos discos não tinha sido abertos"} in \url{https://observador.pt/2021/05/12/diretor-da-pj-confirma-que-rui-pinto-esta-a-colaborar-com-a-justica-e-ja-o-fez-varias-vezes-depois-de-comecar-a-ser-julgado/} },  



% \url{https://desporto.sapo.pt/geral/artigos/football-leaks-rui-pinto-apresenta-queixa-contra-portugal-na-uniao-europeia}
% \url{https://desporto.sapo.pt/geral/artigos/entenda-o-caso-football-leaks-e-o-papel-do-portugues-rui-pinto}
% \url{https://observador.pt/2018/09/14/o-que-se-sabe-de-rui-pinto-o-pirata-informatico-do-football-leaks-que-tera-desviado-os-emails-do-benfica/}
% \url{https://as.com/futbol/2016/03/04/internacional/1457120867_618613.html}